Certified Knowledge Compilation \\ with Application to Verified Model Counting \\ Experimental Details

### 2023-07-25 doc/JAIR/jair-data.tex
Start of JAIR paper

\documentclass[letterpaper,USenglish,cleveref, autoref, thm-restate]{lipics-v2021}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

@@

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Certified Knowledge Compilation \\ with Application to Verified Model Counting}

\titlerunning{Certified Knowledge Compilation}

\author{Randal E. Bryant}{Computer Science Department, Carnegie Mellon University, Pittsburgh, PA 15213 USA}{rebryant@cmu.edu}{https://orcid.org/0000-0001-5024-6613}{Supported by NSF grant CCF-2108521}
\author{Wojciech Nawrocki}{Department of Philosophy, Carnegie Mellon University}{wjnawrock@cmu.edu}{https://orcid.org/0000-0002-8839-0618}{}
\author{Jeremy Avigad}{Department of Philosophy, Carnegie Mellon University}{avigad@cmu.edu}{https://orcid.org/0000-0003-1275-315X}{}
\author{Marijn J. H. Heule}{Computer Science Department, Carnegie Mellon University}{marijn@cmu.edu}{https://orcid.org/0000-0002-5587-8801}{Supported by NSF grant CCF-2108521}

\authorrunning{R. E. Bryant, et al} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

%\Copyright{Jane Open Access and Joan R. Public} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

%\ccsdesc[100]{\textcolor{red}{Replace ccsdesc macro with valid one}} %TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

%\keywords{Dummy keyword} %TODO mandatory; please add comma-separated list of keywords

%\category{} %optional, e.g. invited paper

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

%\nolinenumbers %uncomment to disable line numbering

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\EventEditors{John Q. Open and Joan R. Access}
%\EventNoEds{2}
%\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
%\EventShortTitle{CVIT 2016}
%\EventAcronym{CVIT}
%\EventYear{2016}
%\EventDate{December 24--27, 2016}
%\EventLocation{Little Whinging, United Kingdom}
%\EventLogo{}
%\SeriesVolume{42}
%\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

\section*{Methodology}
\begin{enumerate}
   \item Retrieved 200 public model counting benchmarks from 2022 Model Counting Competition website
     \begin{center}
       \url{https://mccompetition.org/2022/mc_description.html}
     \end{center}
     \begin{itemize}
     \item 100 public instances from Track 1 (unweighted model counting)
     \item 100 public instances from Track 2 (weighted model counting)
     \end{itemize}
   \item Removed duplicates
     \begin{itemize}
    \item  20 instances were duplicated in the two tracks
     \item Reduces set of benchmarks to 180
     \end{itemize}
   \item Ran D4 on 180 instances with time limit of 4000 seconds
     \begin{itemize}
     \item Completed for 124 instances
     \end{itemize}
   \item Generated POG representations from dDNNF graphs
     \begin{itemize}
     \item Simplifications via constant propagation
       \begin{itemize}
       \item Number of POG nodes around ranged between $0.35\times$
         and $1.0\times$ number of dDNNF nodes (harmonic mean =
         $0.81\times$).  Greater optimizations with smaller
         benchmarks.
       \end{itemize}
     \item Used heuristics to identify root node, and then put in topological order with root node at end.
     \item Measure complexity by number of defining clauses $D$
       \begin{itemize}
       \item Product or sum node with degree $k$ requires $k+1$ clauses
       \item $D$ also equals sum of number of edges and number of nodes
       \item For 124 instances, Min = 304, Median = 848,784, Mean = 93,107,937, Max = 2,761,457,765
       \end{itemize}
     \end{itemize}
   \item Ran program {\sc d2p} to generate CPOG file for maximum of
     10,000 seconds.  Verifies file, generates unweighted count. If
     finds literal weights, also generates weighted count
     \begin{itemize}
     \item Completed for 108 instances
     \item Exceeded clause limit for 1 instance (Our programs represent a clause ID as a 32-bit signed value)
     \item Ran out of memory for 1 instance
     \item Timed out for 14 instances
     \item Was able to do one-sided verification for 9 of these.
     \end{itemize}
   \item For completed CPOG files, ran {\sc cpog-check} CPOG checker
     \begin{itemize}
     \item Tests all conditions for correct CPOG
     \item Most time spent performing RUP checks for clause addition and deletion
     \item Completed successfully for all 108 instances
     \item Time generally faster than {\sc d2p}. (min = 0.006, max = 1.38, harmonic mean = 0.107)
     \end{itemize}
\end{enumerate}

>>>>

\documentclass[twoside,11pt]{article}
\usepackage{jair, theapa, rawfonts}

\jairheading{1}{2024}{1--16}{10/22}{03/24}

\ShortHeadings{Certified Knowledge Compilation}{Bryant, Nawrocki, Avigad, \& Heule}

@@

\bibliographystyle{theapa}

\title{Certified Knowledge Compilation \\ with Application to Verified Model Counting \\ Experimental Details }

\author{\name Randal E. Bryant \email rebryant@cmu.edu \\
  \name Wojciech Nawrocki \email wjnawrock@cmu.edu \\
  \name Jeremy Avigad \email avigad@cmu.edu \\
  \name Marijn J. H. Heule \email marijn@cmu.edu \\
  \addr Carnegie Mellon University \\ Pittsburgh, PA  USA}

@@

\begin{abstract}
  This document gives details on the experiments
\end{abstract}

@@

\section*{Methodology}
\begin{enumerate}
   \item Retrieved 200 public model counting benchmarks from 2022 Model Counting Competition website
     \begin{center}
       \url{https://mccompetition.org/2022/mc_description.html}
     \end{center}
     \begin{itemize}
     \item 100 public instances from Track 1 (unweighted model counting)
     \item 100 public instances from Track 2 (weighted model counting)
     \end{itemize}
   \item Removed duplicates
     \begin{itemize}
    \item  20 instances were duplicated in the two tracks.  
     \item Reduced set of benchmarks to 180 by including some from Track 1 and some from Track 2
     \end{itemize}
   \item Ran D4 on 180 instances with time limit of 4000 seconds
     \begin{itemize}
     \item Completed for 123 instances
     \end{itemize}
   \item Generated POG representations from dDNNF graphs
     \begin{itemize}
     \item Simplifications via constant propagation
       \begin{itemize}
       \item Number of POG nodes around ranged between $0.35\times$
         and $1.0\times$ number of dDNNF nodes (harmonic mean =
         $0.81\times$).  Greater optimizations with smaller
         benchmarks.
       \end{itemize}
     \item Used heuristics to identify root node, and then put in topological order with root node at end.
     \item Successfully generated POG representation for 122 benchmarks
       \begin{itemize}
         \item One failed with too many defining clauses (2,761,457,765).  That count exceeded the 32-bit signed integer used to represent clause identifiers.
       \end{itemize}

     \item Measure complexity by number of defining clauses $D$
       \begin{itemize}
       \item Product or sum node with degree $k$ requires $k+1$ clauses
       \item $D$ also equals sum of number of edges and number of nodes
       \item For 122 instances, Min = 304, Median = 848,784, Mean = 93,107,937, Max = 2,761,457,765
       \end{itemize}

     \item Also measure \emph{tree ratio}: $D_t / D$.
       \begin{itemize}
       \item $D_t$ Number of defining clauses if graph were expanded to have no shared subgraphs (i.e., a tree)
       \item $D$ Number of defining clauses for the POG representation
       \item Provides an indication of the importance of exploiting shared subgraphs in the POG generation
       \end{itemize}

     \item Performance (122 instances):
       \begin{itemize}
       \item Times: Min = 0.012 secs, Max = 2804.5 secs, Median = 3.46 secs
       \item Defining clauses: Min = 304, Max = 1,765,743,261, Median = 774,883
       \item Tree ratio: Min = 1.0, Max = 52,409.7, Median = 11.61, Harmonic Mean = 3.21
       \item Tree Sizes ($D_t$): Min = 304, max = 14,924,670,455,484, median = 7,669,582.
       \end{itemize}
     \end{itemize}
\end{enumerate}

\section*{Selecting Reduced Benchmark Set}

To make the experiments tractable, want to select set of benchmarks that are more tractable.
Limit set to the 92 with at most $10^7$ defining clauses.

D4 Performance (92 instances):
       \begin{itemize}
       \item Times: Min = 0.012 secs, Max = 1672.9 secs, Median = 1.85 secs
       \item Defining clauses: Min = 304, Max = 8,493,275, Median = 378,325
       \item Tree ratio: Min = 1.0, Max = 363.9, Median = 9.37, Harmonic Mean = 2.87
       \item Tree Sizes ($D_t$): Min = 304, max = 1,205,743,897, median = 2,690,776
       \end{itemize}

\section*{CPOG Generation}

We have devised two different approaches to generating the forward implication proof from the dec-DNNF generated by {\sc D4}:
\begin{description}
\item[Monolithic:] Make a single call to a proof generating SAT solver.
\item[Structured:] Traverse the POG top-down and recursively generate a forward implication proof for each node encountered.  Lemmas can be introduced to avoid an exponential expansion of the POG into a tree.
\end{description}

Our implementation also allows a hybrid of these approaches: at any
point in the traversal, it can generate a proof for a node via a
single call to a SAT solver.  This capability supports both full
monolithic generation (perform monolithic generation at the POG root),
full structured generation (continue traversing to the POG leaves), or
a balance between the two.
The challenge is then to devise a
heuristic criterion for deciding whether to continue the traversal below
some node or to generate a proof for that node monolithically.

Our first set of experiments applies full monolithic and full structured generation to the reduced benchmark set.  We set a time limit of 1000 seconds for the proof generation.
